Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Paenungulates

### 2020-05-14 paper_PLOS/paper_PLOS_draft.tex
additional figures, cleaned major errors, added completion warnings

\usepackage{longtable}
\usepackage{booktabs}

@@

\emph{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit BLAT
(RBHB) pipeline to quickly identify homologs and estimate gene copy
numbers (\textbf{Figure 1A}). The Reciprocal Best Hit (RBH) search
strategy is conceptually straightforward: 1) Given a gene of interest GA
in a query genome A, one searches a target genome B for all possible
matches to GA; 2) For each of these hits, one then performs the
reciprocal search in the original query genome to identify the
highest-scoring hit; 3) A hit in genome B is defined as a homolog of
gene GA if and only if the original gene GA is the top reciprocal search
hit in genome A. We selected BLAT {[}26{]} as our algorithm of choice,
as this algorithm is sensitive to highly simliar (\textgreater{}90\%
identity) sequences, thus identifying the highest-confidence homologs
while minimizing many-to-one mapping problems when searching for
multiple genes. RBH performs similar to other more complex methods of
orthology prediction, and is particularly good at identifying incomplete
genes that may be fragmented in low quality/poor assembled regions of
the genome {[}{\textbf{???}},27{]}.

\emph{Effective Copy Number By Coverage:} In lower-quality genomes, many
genes are fragmented across multiple scaffolds, which results in BLAT
calling multiple hits when in reality there is only one gene. To
compensate for this, we came up with a novel statistic, Estimated Copy
Number by Coverage (ECNC), which averages the number of times we see
each nucleotides of a query sequence in a target genome over the total
number of nucleotides of the query sequence found overall in each target
genome (\textbf{Figure 1B}; Supplementary Figure 1). This allows us to
correct for genes that have been fragmented across incomplete genomes,
without underestimating the copy number of proteins that lack domains
found in the human sequence.

@@

\emph{Duplication gene inclusion criteria:} In order to condense
transcript-level hits into single gene loci, and to resolve many-to-one
genome mappings, we removed exons where transcripts from different genes
overlapped, and merged overlapping transcripts of the same gene into a
single gene locus call. The resulting gene-level copy number table was
then combined with the maximum ECNC values observed for each gene in
order to call gene duplications. We called a gene duplicated if its copy
number was two or more, and if the maximum ECNC value of all the gene
transcripts searched was 1.5 or greater; previous studies have shown
that incomplete duplications can encode functional genes, therefore
partial gene duplications were included provided they passed additional
inclusion criteria. The ECNC cut off of 1.5 was selected empirically, as
this value minimized the number of false positives seen in a test set of
genes and genomes. The results of our initial search are summarized in
Figure 1B. Overall, we identified {[}MEDIAN{]} genes across all species,
or {[}\%HITS/QUERIES{]} of our starting query genes. As expected, there
was a correlation between the number of hits identified in a species
with the evolutionary distance relative to humans {[}Sup Figure 1{]}.

@@

@@

To determine which pathways were associated with duplicated genes in
each species and lineage, we used WEBGESTALT to perform
overrepresentation analysis (ORA) of the duplicated gene lists relative
to our initial query gene list. For the database of pathways used in the
analysis, we used Reactome {[}{\textbf{???}}{]}, Wikipathways and
Wikipathways\_cancer {[}{\textbf{???}}{]}, and KEGG
{[}{\textbf{???}}{]}.

@@

\begin{verbatim}
## Warning: Tried to calculate with group_by(), but the calculation failed.
## Falling back to ungrouped filter operation...

## Warning: Tried to calculate with group_by(), but the calculation failed.
## Falling back to ungrouped filter operation...
\end{verbatim}

\begin{verbatim}
## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
##   If you'd like to see this geom implemented,
##   Please open an issue with your example code at
##   https://github.com/ropensci/plotly/issues

## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
##   If you'd like to see this geom implemented,
##   Please open an issue with your example code at
##   https://github.com/ropensci/plotly/issues

## Warning in geom2trace.default(dots[[1L]][[3L]], dots[[2L]][[1L]], dots[[3L]][[1L]]): geom_GeomTextRepel() has yet to be implemented in plotly.
##   If you'd like to see this geom implemented,
##   Please open an issue with your example code at
##   https://github.com/ropensci/plotly/issues
\end{verbatim}

\hypertarget{step-wise-evolution-of-large-long-lived-afrotherians}{%
\subsection{Step-wise evolution of large, long-lived
Afrotherians}\label{step-wise-evolution-of-large-long-lived-afrotherians}}

\begin{verbatim}
## Warning: Removed 120 rows containing missing values (geom_image).
\end{verbatim}

@@

\begin{verbatim}
## Warning: Column `Genome` joining factor and character vector, coercing into
## character vector
\end{verbatim}

\begin{verbatim}
## Warning in melt(., na.rm = T): The melt generic in data.table has been passed a
## matrix and will attempt to redirect to the relevant reshape2 method; please note
## that reshape2 is deprecated, and this redirection is now deprecated as well.
## To continue using melt methods from reshape2 while both libraries are attached,
## e.g. melt.list, you can prepend the namespace like reshape2::melt(.). In the
## next version, this warning will become an error.
\end{verbatim}

\begin{verbatim}
##                                        Var1 Var2        value
## 1                                      Gene Copy  0.011177676
## 2                                    Genome Copy -0.104365173
## 3       Sum length of sequences > 10M  (nt) Copy -0.074960529
## 4                     Longest sequence (nt) Copy -0.067231077
## 5                  N50 sequence length (nt) Copy -0.050741921
## 6                  Base composition (%) (A) Copy -0.001502967
## 7                  Base composition (%) (T) Copy -0.001379093
## 8                  Base composition (%) (G) Copy -0.043271156
## 9                  Base composition (%) (C) Copy -0.042268177
## 10             Base composition (%) (Other) Copy -0.050228064
## 11 # of sequences containing non-ACGTN (nt) Copy -0.050621596
## 12                Mean sequence length (nt) Copy -0.074983744
## 13              Median sequence length (nt) Copy -0.074897040
## 14                         Missing Proteins Copy -0.079692396
## 15                                   #Prots Copy -0.101300994
## 16                             Completeness Copy -0.101321644
## 17                 Base composition (%) (N) Copy  0.019768158
## 18                              totalLength Copy  0.067454870
## 19       Sum length of sequences > 1M  (nt) Copy  0.051354838
## 20               # of sequences > 100K (nt) Copy  0.056049593
## 21                                      L50 Copy  0.082175611
## 22               # of sequences >   1M (nt) Copy  0.074461957
## 23               # of sequences >  10M (nt) Copy  0.075679451
## 24                   Shortest sequence (nt) Copy -0.013216232
## 25                           GC-content (%) Copy -0.064591581
## 26                           # of sequences Copy  0.076630197
## 27               # of sequences >   1K (nt) Copy  0.047474518
## 28               # of sequences >  10K (nt) Copy  0.045583001
## 29                                   #Total Copy  0.116391876
## 30                                  Average Copy  0.124341009
## 31                                   %Ortho Copy  0.120616153
## 32                                     Gene ECNC  0.011137324
## 33                                   Genome ECNC -0.090522003
## 34      Sum length of sequences > 10M  (nt) ECNC -0.061380983
## 35                    Longest sequence (nt) ECNC -0.053145168
## 36                 N50 sequence length (nt) ECNC -0.040007282
## 37                 Base composition (%) (A) ECNC  0.004626621
## 38                 Base composition (%) (T) ECNC  0.004862588
## 39                 Base composition (%) (G) ECNC -0.037787699
## 40                 Base composition (%) (C) ECNC -0.037045693
## 41             Base composition (%) (Other) ECNC -0.040274294
## 42 # of sequences containing non-ACGTN (nt) ECNC -0.040587311
## 43                Mean sequence length (nt) ECNC -0.061569667
## 44              Median sequence length (nt) ECNC -0.060928944
## 45                         Missing Proteins ECNC -0.062209031
## 46                                   #Prots ECNC -0.084046941
## 47                             Completeness ECNC -0.084067608
## 48                 Base composition (%) (N) ECNC  0.013486259
## 49                              totalLength ECNC  0.061175951
## 50       Sum length of sequences > 1M  (nt) ECNC  0.046387082
## 51               # of sequences > 100K (nt) ECNC  0.049326449
## 52                                      L50 ECNC  0.070430706
## 53               # of sequences >   1M (nt) ECNC  0.063725329
## 54               # of sequences >  10M (nt) ECNC  0.062895861
## 55                   Shortest sequence (nt) ECNC -0.012113794
## 56                           GC-content (%) ECNC -0.066249117
## 57                           # of sequences ECNC  0.070289641
## 58               # of sequences >   1K (nt) ECNC  0.042151677
## 59               # of sequences >  10K (nt) ECNC  0.041222984
## 60                                   #Total ECNC  0.102797584
## 61                                  Average ECNC  0.108543367
## 62                                   %Ortho ECNC  0.104394790
## 63                                     Gene   CN  0.013089940
## 64                                   Genome   CN -0.092501845
## 65      Sum length of sequences > 10M  (nt)   CN -0.061610123
## 66                    Longest sequence (nt)   CN -0.052591263
## 67                 N50 sequence length (nt)   CN -0.038770715
## 68                 Base composition (%) (A)   CN  0.006139187
## 69                 Base composition (%) (T)   CN  0.006354445
## 70                 Base composition (%) (G)   CN -0.036741673
## 71                 Base composition (%) (C)   CN -0.035971303
## 72             Base composition (%) (Other)   CN -0.041544748
## 73 # of sequences containing non-ACGTN (nt)   CN -0.041868348
## 74                Mean sequence length (nt)   CN -0.063786387
## 75              Median sequence length (nt)   CN -0.063541494
## 76                         Missing Proteins   CN -0.063703014
## 77                                   #Prots   CN -0.085286259
## 78                             Completeness   CN -0.085306203
## 79                 Base composition (%) (N)   CN  0.012122309
## 80                              totalLength   CN  0.060342667
## 81       Sum length of sequences > 1M  (nt)   CN  0.044889905
## 82               # of sequences > 100K (nt)   CN  0.047874628
## 83                                      L50   CN  0.069838784
## 84               # of sequences >   1M (nt)   CN  0.062839574
## 85               # of sequences >  10M (nt)   CN  0.062215472
## 86                   Shortest sequence (nt)   CN -0.013168670
## 87                           GC-content (%)   CN -0.066875760
## 88                           # of sequences   CN  0.073563144
## 89               # of sequences >   1K (nt)   CN  0.044231140
## 90               # of sequences >  10K (nt)   CN  0.042436480
## 91                                   #Total   CN  0.105113732
## 92                                  Average   CN  0.110764419
## 93                                   %Ortho   CN  0.106456398
\end{verbatim}

\includegraphics{paper_PLOS_draft_files/figure-latex/Correlations between genome quality scores-1.pdf}

\begin{verbatim}
## Warning: Column `Genome`/`BestGenome` joining factor and character vector,
## coercing into character vector
\end{verbatim}

\begin{verbatim}
## Warning: `as.tibble()` is deprecated as of tibble 2.0.0.
## Please use `as_tibble()` instead.
## The signature and semantics have changed, see `?as_tibble`.
## This warning is displayed once every 8 hours.
## Call `lifecycle::last_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{figure}
\centering
\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2-1.pdf}
\caption{Figure 2: A Reciprocal Best-Hit BLAT pipeline for identifying
gene copy number in other genomes. A) A graphic summary of the
reciprocal best-hit strategy. B) Estimated Copy Number by Coverage. C)}
\end{figure}

\begin{verbatim}
## Warning: Table 2
\end{verbatim}

Enhanced cancer suppression may have evolved through many mechanisms;
among the most parsimonious is an increase in the copy number of genes
with tumor suppressor functions. Previous studies focusing on candidate
gene studies, for example, have identified increased copy number of the
tumor suppressors \emph{TP53} and \emph{LIF} in elephants
{[}{\textbf{???}},11,21--23{]}. Therefore, in order to test whether this
was a pervasive phenomena genome-wide in \emph{Afrotherians}, we used a
Reciprocal Best Hit BLAT (RBHB) approach to infer gene copy number in
\emph{Afrotherian} and \emph{Atlantogenatan} genomes (Fig. 2A). Because
RBHB-like approaches can over-estimate copy number when genes are
fragmented or incorrectly assembled across multiple scaffolds, we also
inferred copy number using a complementary method that quantifies the
ratio between observed and expected gene coverage per nucleotide (ECNC)
(Fig. 2B; Sup. Fig. 2A-D). By only including nucleotides from the query
sequence that were observed in the target genome, we also correct for
partial hits where some or all of the homologs of a gene have diverged
from the human homolog.

@@

@@

\hypertarget{retroduplication-of-eef1a1-in-the-stem-lineage-of-tehthytheria}{%
\subsection{Retroduplication of EEF1A1 in the stem-lineage of
Tehthytheria}\label{retroduplication-of-eef1a1-in-the-stem-lineage-of-tehthytheria}}

\begin{verbatim}
## Warning: Fig 3: EEF1A1
\end{verbatim}

Previous studies have identified genes whose transgenic overexpression
delays aging related phenotypes and extends lifespan. Overexpression of
the eukaryotic translation elongation factor 1 alpha 1 (EEF1A1) gene,
for example, extends lifespan in transgenic Drosophila. Like most
mammalian genomes, we found that Afrotherian genomes encoded numerous
EEF1A1 retrogenes (Fig. XA). While the majority of these retrogenes have
premature stop codons and insertions/deletions characteristic of
pseudogenes, at least 13 Tehthytherian-specific EEF1A1 retrogenes have
the potential to encode functional proteins (Fig. XB). Unfortunately it
is difficult to determine if these retrogenes are transcribed because
they have relatively high sequence similarity (83-88\%), suggesting
evidence of transcription using RNA-Seq data is likely an artifact.
Indeed, nearly all of the EEF1A1 retrogene transcripts assembled by
StringTie were confined to their gene body and UTRs and thus lacked
unique sequences. The EEF1A1RTG13 transcript, however, initiates within
an upstream MLT1B retrovirus-like MaLR transposable element and includes
unique 5?- and 3?-UTR sequences, indicating at least one EEF1A1
retrogene is transcribed (Fig. XC). EEF1A1

Candidate gene studies, for example, have identified functional
duplicates of the tumor suppressors TP53 and LIF in elephants. In a
larger candidate gene study, Caulin et al.~characterized the copy number
of 830 known tumor-suppressor genes across 36 mammals and identified 382
putative duplicates, including duplicates in species with large body
sizes and long life-spans. however, the probability of developing cancer
is similar for small, short-lived mammals such as mice and for large,
long-lived mammals such as elephants.

@@

\begin{verbatim}
## Warning: Fig4
\end{verbatim}

\begin{verbatim}
## # A tibble: 253 x 6
##     parent  node branch.length label                       lnSize Lifespan
##      <int> <int>         <dbl> <chr>                        <dbl>    <dbl>
##   1    135     1     40.1      Geogale.aurita                   2    1.72 
##   2    140     2    122.       Limnogale.mergulus               5    2.30 
##   3    140     3      0.135    Microgale.taiva                  2    1.72 
##   4    139     4     33.5      Microgale.parvula                1    1.52 
##   5    138     5      0.595    Microgale.brevicaudata           2    1.72 
##   6    142     6      0.571    Microgale.cowani                 3    1.91 
##   7    142     7      1.07     Microgale.jobihely               2    1.72 
##   8    145     8      0.772    Microgale.longicaudata           2    1.72 
##   9    145     9      0.173    Microgale.principula             2    1.72 
##  10    144    10     34.3      Microgale.pusilla                1    1.52 
##  11    143    11     61.4      Microgale.talazaci               4    1.76 
##  12    136    12     22.8      Oryzorictes.hova                 4    2.11 
##  13    147    13      7.75     Echinops.telfairi                4    2.94 

>>>>

\usepackage{longtable}
\usepackage{booktabs}
\usepackage{geometry}
\usepackage{pdflscape}

@@

\emph{Reciprocal Best-Hit BLAT:} We developed a reciprocal best hit BLAT
(RBHB) pipeline to quickly identify homologs and estimate gene copy
numbers (\textbf{Figure 1A}). The Reciprocal Best Hit (RBH) search
strategy is conceptually straightforward: 1) Given a gene of interest
\(G_A\) in a query genome \(A\), one searches a target genome \(B\) for
all possible matches to \(G_A\); 2) For each of these hits, one then
performs the reciprocal search in the original query genome to identify
the highest-scoring hit; 3) A hit in genome \(B\) is defined as a
homolog of gene \(G_A\) if and only if the original gene \(G_A\) is the
top reciprocal search hit in genome \(A\). We selected BLAT {[}26{]} as
our algorithm of choice, as this algorithm is sensitive to highly
simliar (\textgreater{}90\% identity) sequences, thus identifying the
highest-confidence homologs while minimizing many-to-one mapping
problems when searching for multiple genes. RBH performs similar to
other more complex methods of orthology prediction, and is particularly
good at identifying incomplete genes that may be fragmented in low
quality/poor assembled regions of the genome {[}{\textbf{???}},27{]}.

\emph{Effective Copy Number By Coverage:} In lower-quality genomes, many
genes are fragmented across multiple scaffolds, which results in BLAT
calling multiple hits when in reality there is only one gene. To
compensate for this, we came up with a novel statistic, Estimated Copy
Number by Coverage (ECNC), which averages the number of times we see
each nucleotides of a query sequence in a target genome over the total
number of nucleotides of the query sequence found overall in each target
genome (Supplementary Figure 1). This allows us to correct for genes
that have been fragmented across incomplete genomes, while also taking
into account missing sequences from the human query in the target
genome. Mathematically, this can be written as:\\
\[ ECNC = \frac{\sum_{n=1}^{l} C_n}{\sum_{n=1}^{l} bool(C_n)}\] where
\(n\) is a given nucleotide in the query, \(l\) is the total length of
the query, \(C_n\) is the number of instances that \(n\) is present
within a reciprocal best hit, and \(bool(C_n)\) is 1 if \(C_n > 0\) or 0
if \(C_n = 0\).

@@

\emph{Duplication gene inclusion criteria:} In order to condense
transcript-level hits into single gene loci, and to resolve many-to-one
genome mappings, we removed exons where transcripts from different genes
overlapped, and merged overlapping transcripts of the same gene into a
single gene locus call. The resulting gene-level copy number table was
then combined with the maximum ECNC values observed for each gene in
order to call gene duplications. We called a gene duplicated if its copy
number was two or more, and if the maximum ECNC value of all the gene
transcripts searched was 1.5 or greater; previous studies have shown
that incomplete duplications can encode functional genes, therefore
partial gene duplications were included provided they passed additional
inclusion criteria. The ECNC cut off of 1.5 was selected empirically, as
this value minimized the number of false positives seen in a test set of
genes and genomes. The results of our initial search are summarized in
Figure 1B. Overall, we identified {[}MEDIAN{]} genes across all species,
or {[}\%HITS/QUERIES{]} of our starting query genes.

@@

\textbf{Orthogonal Genome Assessment using CEGMA} In order to determine
the effect of genome quality on our results, we used the gVolante
webserver and CEGMA to assess the quality and completeness of the
genome. CEGMA was run using the default settings of {[}{]}, and the
mammalian-specific core gene sets.

@@

To determine which pathways were associated with duplicated genes in
each species and lineage, we used WEBGESTALT to perform
overrepresentation analysis (ORA) of the duplicated gene lists relative
to our initial query gene list {[}{\textbf{???}}{]}. For the database of
pathways used in the analysis, we used Reactome {[}{\textbf{???}}{]},
Wikipathways, and Wikipathways\_cancer {[}{\textbf{???}}{]}, and KEGG
{[}{\textbf{???}}{]}. For the ORA, we used FDR for determining
significance, and ran the analysis at FDR=0.1, FDR=0.2, FDR=0.3, and
FDR=0.5.

\hypertarget{lifespan-phylogenetic-least-square-regression-and-calculating-estimated-cancer-risk-thoughout-atlantogenata}{%
\subsection{Lifespan Phylogenetic Least-Square Regression and
Calculating Estimated Cancer Risk Thoughout
Atlantogenata}\label{lifespan-phylogenetic-least-square-regression-and-calculating-estimated-cancer-risk-thoughout-atlantogenata}}

In order to determine the cancer risk at each node, we used a simplified
multistage cancer risk model for body size and lifespan
{[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. We defined the

@@

\hypertarget{body-size-frequently-and-independently-expands-and-contracts-throughout-atlantogenata}{%
\subsection{\texorpdfstring{Body size frequently and independently
expands and contracts throughout
\emph{Atlantogenata}}{Body size frequently and independently expands and contracts throughout Atlantogenata}}\label{body-size-frequently-and-independently-expands-and-contracts-throughout-atlantogenata}}

\begin{figure}
\centering
\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 1-1.pdf}
\caption{Figure 1: Body sizes rapidly and frequently expand in
Eutherians, especially in Atlantogenata. \textbf{A)} Tree of Eutherian
species, colored by ln(Body Size) and with branch lengths set to the
rate of change in body sizes, normalized by the square root of the root
branch. Atlantogenata is highlighted at the bottom. \textbf{B)} Zoom-in
of (A) on Atlantogenata. Silhuetes for the African Elephant, West Indian
Manatee, Cape Elephant Shrew, Lesser Hedgehog Tenrec, Cape Golden Mole,
Nine-Banded Armadillo, and Hoffman's Two-Toed Sloth are colored by their
extant body sizes, while clade labels are colored based on the common
ancestor's estimated body size}
\end{figure}

@@

\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2C1-1.pdf}
\includegraphics{paper_PLOS_draft_files/figure-latex/Figure 2C2-1.pdf}

\begin{verbatim}
## Warning: TODO: Table 2 is done, but Stargazer has some weird bug stopping it
## from running...
\end{verbatim}

Enhanced cancer suppression may have evolved through many mechanisms;
among the most parsimonious is an increase in the copy number of genes
with tumor suppressor functions. Previous studies focusing on candidate
gene studies, for example, have identified increased copy number of the
tumor suppressors \emph{TP53} and \emph{LIF} in elephants
{[}{\textbf{???}},11,21--23{]}. Therefore, in order to test whether this
was a pervasive phenomena genome-wide in \emph{Afrotherians}, we used a
Reciprocal Best Hit BLAT (RBHB) approach to infer gene copy number in
\emph{Afrotherian} and \emph{Atlantogenatan} genomes (Fig. 2A). Because
RBHB-like approaches can over-estimate copy number when genes are
fragmented or incorrectly assembled across multiple scaffolds, we also
inferred copy number using a complementary method that quantifies the
ratio between observed and expected gene coverage per nucleotide (ECNC)
(Supplementary Figure 1). By only including nucleotides from the query
sequence that were observed in the target genome, we also correct for
partial hits where some or all of the homologs of a gene have diverged
from the human homolog.

@@

\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 3-1.pdf}
\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 3-2.pdf}

@@

\begin{verbatim}
## Warning: TODO: Make Fig4
\end{verbatim}

\begin{verbatim}
## Warning: TODO: Caption Fig 5
\end{verbatim}

\begin{verbatim}
## Warning: TODO: de-comment geom_tiplab in f5a! Commented currently for offline
## usage!
\end{verbatim}

\includegraphics{paper_PLOS_draft_files/figure-latex/Fig 5-1.pdf}

\begin{verbatim}
## 
## % Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
## % Date and time: Wed, May 13, 2020 - 20:08:58
## \begin{table}[!htbp] \centering 
##   \caption{Phylogenetic Least Squares: ln(Lifespan) & ln(Body Size) Regression} 
##   \label{} 
## \begin{tabular}{@{\extracolsep{5pt}}lc} 
## \\[-1.8ex]\hline 
## \hline \\[-1.8ex] 
##  & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
## \cline{2-2} 
## \\[-1.8ex] & Lifespan \\ 
## \hline \\[-1.8ex] 
##  lnSize & 0.100 \\ 
##   & (0.121) \\ 
##   & t = 0.826 \\ 
##   & p = 0.409 \\ 
##   Constant & 1.943 \\ 
##   & (1.385) \\ 
##   & t = 1.403 \\ 
##   & p = 0.161 \\ 
##  \hline \\[-1.8ex] 
## Observations & 28 \\ 
## Log Likelihood & $-$39.636 \\ 
## Akaike Inf. Crit. & 85.273 \\ 
## Bayesian Inf. Crit. & 89.047 \\ 
## \hline 
## \hline \\[-1.8ex] 
## \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
## \end{tabular} 
## \end{table}
\end{verbatim}

\begin{longtable}[]{@{}lrrllrrl@{}}
\caption{A table generated by the longtable package.}\tabularnewline
\toprule
Node & lnSize & Est. Lifespan & Est. Cancer Susceptibility (K1) & Est.
Ancestral K2 & Change in K (K2/K1) & log2 Fold Change &
Ancestor\tabularnewline
\midrule
\endfirsthead
\toprule
Node & lnSize & Est. Lifespan & Est. Cancer Susceptibility (K1) & Est.
Ancestral K2 & Change in K (K2/K1) & log2 Fold Change &
Ancestor\tabularnewline
\midrule
\endhead
Loxodontini & 16 & 34.38 & 1.47e+16 & 2.97e+15 & 4.940000e+00 & 2.31 &
Elephantidae\tabularnewline
Loxodonta africana & 15 & 65.00 & 2.47e+17 & 1.47e+16 & 1.681000e+01 &
4.07 & Loxodontini\tabularnewline
Loxodona & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 & 0.00 &
Loxodontini\tabularnewline
Loxodonta cyclotis & 15 & 31.12 & 2.97e+15 & 1.47e+16 & 2.000000e-01 &
-2.31 & Loxodona\tabularnewline
Palaeoloxodon antiquus & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00
& 0.00 & Loxodona\tabularnewline
Elephantidae & 15 & 31.12 & 2.97e+15 & 1.40e+07 & 2.125534e+08 & 27.66 &
NA\tabularnewline
Elephantina & 16 & 34.38 & 1.47e+16 & 2.97e+15 & 4.940000e+00 & 2.31 &
Elephantidae\tabularnewline
Elephas maximus & 15 & 65.50 & 2.58e+17 & 1.47e+16 & 1.760000e+01 & 4.14
& Elephantina\tabularnewline
Mammuthus & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 & 0.00 &
Elephantina\tabularnewline
Mammuthus primigenius & 15 & 31.12 & 2.97e+15 & 1.47e+16 & 2.000000e-01
& -2.31 & Mammuthus\tabularnewline
Mammuthus columbi & 16 & 34.38 & 1.47e+16 & 1.47e+16 & 1.000000e+00 &
0.00 & Mammuthus\tabularnewline
Mammut americanum & 16 & 34.38 & 1.47e+16 & 1.40e+07 & 1.050567e+09 &
29.97 & NA\tabularnewline
Tethytheria & 13 & 25.49 & 1.21e+14 & 1.01e+12 & 1.207400e+02 & 6.92 &
Paenungulata\tabularnewline
Trichechus manatus & 13 & 69.00 & 4.77e+16 & 1.21e+14 & 3.929800e+02 &
8.62 & Tethytheria\tabularnewline
Paenungulata & 10 & 18.91 & 1.01e+12 & 1.01e+12 & 1.000000e+00 & 0.00 &
Pseudoungulata\tabularnewline
Procavia capensis & 8 & 14.80 & 3.13e+10 & 1.01e+12 & 3.000000e-02 &
-5.01 & Paenungulata\tabularnewline
Pseudoungulata & 10 & 18.91 & 1.01e+12 & 1.69e+09 & 5.967900e+02 & 9.22
& Afroinsectivora\tabularnewline
Orycteropus afer & 11 & 29.80 & 4.19e+13 & 1.01e+12 & 4.167000e+01 &
5.38 & Pseudoungulata\tabularnewline
Elephantulus edwardii & 4 & 10.40 & 6.90e+07 & 1.69e+09 & 4.000000e-02 &
-4.61 & Afroinsectivora\tabularnewline
Afrosorcida & 4 & 10.40 & 6.90e+07 & 1.69e+09 & 4.000000e-02 & -4.61 &
Afrotheria\tabularnewline
Chrysochloris asiatica & 4 & 10.40 & 6.90e+07 & 6.90e+07 & 1.000000e+00
& 0.00 & NA\tabularnewline
Echinops telfairi & 4 & 19.00 & 2.57e+09 & 6.90e+07 & 3.722000e+01 &
5.22 & NA\tabularnewline
Afrotheria & 6 & 12.69 & 1.69e+09 & 2.83e+06 & 5.967900e+02 & 9.22 &
Atlantogenata\tabularnewline
Xenarthra & 11 & 20.89 & 4.97e+12 & 2.83e+06 & 1.760358e+06 & 20.75 &
Atlantogenata\tabularnewline
Dasypus novemcinctus & 8 & 22.30 & 3.67e+11 & 4.97e+12 & 7.000000e-02 &
-3.76 & Xenarthra\tabularnewline
Choloepus hoffmanni & 8 & 41.00 & 1.42e+13 & 4.97e+12 & 2.850000e+00 &
1.51 & Xenarthra\tabularnewline
Atlantogenata & 2 & 8.52 & 2.83e+06 & 2.83e+06 & 1.000000e+00 & 0.00 &
Atlantogenata\tabularnewline
Afroinsectivora & 6 & 12.69 & 1.69e+09 & 1.69e+09 & 1.000000e+00 & 0.00
& Afrotheria\tabularnewline
NA & 3 & 9.41 & 1.40e+07 & 1.21e+14 & 0.000000e+00 & -23.05 &
Tethytheria\tabularnewline
\bottomrule
\end{longtable}

The dramatic increase in body mass and lifespan in some Afrotherian
lineages implies those lineages evolved reduced cancer risk. To infer
the magnitude of these reductions we estimated differences in cancer
risk between small bodied, short-lived species and large bodied,
long-lived species as well as for reconstructed ancestral Afrotherians.
Following {[}{\textbf{???}}{]} we estimate the intrinsic cancer risk as
the product of risk associated with body mass and lifespan. Differences
in cancer susceptibility \(K\) due to body mass differences between
species can be approximated simply as the fold difference in body mass
(\(D\)) between species {[}{\textbf{???}}{]}. The risk of developing
cancer also increases in proportion to the sixth power of age and is
approximated by the formula \(Ct^6\), in which the proportionality
constant C that determines susceptibility to cancer induction is
multiplied by the sixth power of the age in years, \(t\)
{[}{\textbf{???}},{\textbf{??}},{\textbf{??}}{]}. Thus we can estimate
the intrinsic cancer risk for a species as \(K \approx Dt^6\).

In order to estimate the intrinsic cancer risk of a species, we first
obtained estimates for lifespans at ancestral nodes using PGLS and the
model
\(ln(lifespan) = \beta_{1}corBrownian +\beta_{2}ln(Size) + \epsilon\)
(Figure ). With this information in hand, we calculated \(K_{1}\) at all
nodes, and then estimated the fold change in cancer succeptibility
between an ancestral node and a given node as \(\frac{K_{2}}{K_{1}}\)
(\textbf{Table 4}).

As shown in \textbf{Table 4}, cancer succeptibility skyrocketed at the
initial divergence of Atlantogenata, followed by a generally upwards
trend. At the common ancestor of Afrotheria there is an inital 9.22-fold
increase in cancer risk. In parallel to Afrotheria, cancer
succeptibility increases 20.75-fold in Xenarthra. However, cancer risk
slowly deflates as size decreases as one moves along the tree towards
extant species, such as in Hoffman's Two Toed Sloth (-fold change) and
in the Nine-banded Armadillo (-fold change).

Within Afrotheria, cancer succeptibility drops in Afrosorcida as species
shrink (-4.61-fold, then stagnates for the Cape Golden Mole) - but then
rises -fold towards the Lesser Hedgehog Tenrec. In parallel,
Afroinsectivora does not increase in cancer succeptibility, and
decreases once more at the Cape Elephant Shrew (-fold). The emergence of
Pseudoungulata sees the next big leap in cancer succeptibility with a
9.22-fold increase. The Aardvark further increases -fold, while we don't
observe an increase at the common ancestor of Paenungulates. While the
Rock Hyrax decreases in cancer succeptibility as expected (-fold),
Tethytheria sees a sharp increase in cancer risk (6.92-fold). Within
Tethytheria, the Manatee's cancer risk increases once more -fold, while
Proboscidae's cancer risk drops precipitously along with its body size
(-23.05-fold). Yet, within Proboscidae we see the biggest increases:
right off the bat, we see that the cancer succeptibility of Elephantidae
and the American Mastodon skyrocket by 27.66-fold and -fold,
respectively. Both Elephantina and Loxodontini in Elephantidae have a
2.31-fold increase in cancer succeptibility. Within Elephantina, cancer
susceptibility stays stable at Mammuthus and in the Colombian Mammoth,
and slightly decreases in the Wooly Mammoth (-fold). The three extant
elephants - Asian Elephant in Elephantina, the African Savana Elephant
in Loxodontini, and the African Forest Elephant in Loxodona, meanwhile,
have parallel and similar decreases in both size and cancer
succeptibility (-, -, and -fold, respectively). Neither the common
ancestor of Loxodonta, nor the Straight-Tusked Mammoth see any further
changes in cancer succeptibility.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

What biological mechanisms underlie the evolution of thousand- to
hundred million- increases in cancer susceptibility during the origins
of Afrotherians, which are essential for large body size and long
lifespan to evolve

Candidate gene studies, for example, have identified functional
duplicates of the tumor suppressors TP53 and LIF in elephants. In a
larger candidate gene study, Caulin et al.~characterized the copy number
of 830 known tumor-suppressor genes across 36 mammals and identified 382
putative duplicates, including duplicates in species with large body
sizes and long life-spans. However, the probability of developing cancer
is similar for small, short-lived mammals such as mice and for large,
long-lived mammals such as elephants.

@@

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

I would like to thank Olga Duchenko at the Aiden Lab, D.H. Vazquez for
his indispensible support.

\hypertarget{conflicts-of-interest}{%
\section{Conflicts of Interest}\label{conflicts-of-interest}}

The Authors have no conflicts of interest to report

\hypertarget{funding-source}{%
\section{Funding Source}\label{funding-source}}

We would like to thank the Department of Human Genetics at the
University of Chicago for supporting this project.

\hypertarget{supplementary-figures}{%
\section{SUPPLEMENTARY FIGURES}\label{supplementary-figures}}
